---
title: "Untitled"
author: "YOU"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
###It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

###This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


```{r echo=TRUE}
unzip("./activity.zip")
activity<-read.csv("./activity.csv",header = TRUE)
activity[,2]<-as.Date(activity$date)
str(activity)
```
```{r echo=TRUE}
steps_1<-with(activity,tapply(steps,date,sum,na.rm=TRUE))
hist(steps_1,col = "green",xlab = "Total Steps",ylab = "Frequency",main = "Total Number of Steps per Day")
```
### The average total steps and median of total steps taken each day
```{r echo=TRUE}
print(mean_steps<-mean(steps_1))
```
```{r echo=TRUE}
print(median_steps<-median(steps_1))
```
### So the average total steps taken each day is 9354.2295082 and, the average median is is calculated to be 10395
###The summary of the total steps taken each day

```{r echo=TRUE}
summary(steps_1)
```
## **What is the average daily activity pattern?**

### a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r echo=TRUE}
avg_steps<-with(activity,tapply(steps,interval,mean,na.rm=TRUE))
intervals<-unique(activity$interval)
new<-data.frame(cbind(avg_steps,intervals))
plot(new$intervals,new$avg_steps,type = "l", xlab = "Intervals",
     ylab = "Average Steps",main = "Average Steps per Interval")
```
### which 5-minute interval contains the maximum number of steps
```{r echo=TRUE}
index<-which.max(new$avg_steps)
max<-new[index,2]
```
###5-minute interval that contains the maximum number of steps is 835

## **Imputing missing values**

### Approach: Calculate the average of average steps per day across all dates in the data set (ignoring NA values). Then use the resulting value in place of NAs.
```{r echo=TRUE}
sum(is.na(activity$steps))
```
```{r echo=TRUE}
steps_2<-with(activity,tapply(steps,date,sum,na.rm=TRUE))
hist(steps_2,col = "green",xlab = "Total Steps",ylab = "Frequency",main = "Total Number of Steps per Day")
```
### the mean and median of the total steps taken each day
```{r echo=TRUE}
print(mean_steps_2<-mean(steps_2))
```
```{r echo=TRUE}
print(mean_steps_2<-mean(steps_2))
```
### The average total steps taken each day is 1.076618910^{4} and the median of total steps taken each day is 1.076618910^{4}. Notice that both the average and the median of the total steps taken eah day after NAs are filled came out to be equal.
 
## **Are there differences in activity patterns between weekdays and weekends?**

```{r echo=TRUE}
library(dplyr)
```

```{r echo=TRUE}
activity_mod<- mutate(activity, day = ifelse(weekdays(activity$date) == "Saturday" | weekdays(activity$date) == "Sunday", "weekend", "weekday"))
activity_mod$day<-as.factor(activity_mod$day)
str(activity_mod)
```

```{r echo=TRUE}
act_wknd<-subset(activity_mod,as.character(activity_mod$day)=="weekend")
act_wkdy<-subset(activity_mod,as.character(activity_mod$day)=="weekday")
steps_wknd<-with(act_wknd,tapply(steps,interval,mean,na.rm=TRUE))
steps_wkdy<-with(act_wkdy,tapply(steps,interval,mean,na.rm=TRUE))
int_wknd<-unique(act_wknd$interval)
int_wkdy<-unique(act_wkdy$interval)
new_wknd<-data.frame(cbind(steps_wknd,int_wknd))
new_wkdy<-data.frame(cbind(steps_wkdy,int_wkdy))
par(mfrow=c(2,1),mar=c(4,4,2,1))
plot(new_wknd$int_wknd,new_wknd$steps_wknd,type = "l",xlab = "Intervals",
     ylab = "Average Steps",main = "Weekend")
plot(new_wkdy$int_wkdy,new_wkdy$steps_wkdy,type = "l",xlab = "Intervals",
     ylab = "Average Steps",main = "Weekday")
```











